Foreground Signatures and Mitigation in Neutral Hydrogen Power Spectra From The Reionization Epoch

### 2014-10-23 foreground_delay_spectrum_modeling.tex
from 449b860: ============
(*) Incorporated Judd's comments
(*) Changed a few figures

Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,mor12,tro12,thy13,liu14a,liu14b}. Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity. Foregrounds can be described in both the delay spectrum and Fourier frameworks. For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite their subtle differences.

@@

Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper. $|\vec{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis. $\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge}, whose boundaries are determined by antenna spacing and the light travel time across it. These boundaries are called horizon delay limits and are shown by solid lines. The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis. This effect has been quantified in \citet{thy13}. The width of this narrow strip is inversely proportional to the operating bandwidth. The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades. However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s. The underlying cause of this effect will be discussed later in this paper.

@@

We show results of delay spectrum from MWA data after deconvolution along the delay axis in Figure~\ref{fig:fhd_data}. The principle feature is the {\it foreground wedge}, whose boundaries, marked by solid white lines, increase linearly with light travel time corresponding to baseline length. Note that the wedge is not filled out uniformly. Each baseline has a different orientation and resulting response to different objects in the sky. It is this variation that we will explore below.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{Delay spectra of two observations of the EoR field (RA = 0h, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set. To get the delay spectrum, each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted in order of baseline length ($x$--axis). For reference, the $k$--space equivalent of $\tau$ and $|\vec{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. The logarithmic color scale (shown at the right) is common to both panels. White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline. Smooth spectrum objects predominantly reside in this {\it foreground wedge}. The various features in this region depend on antenna spacing orientation, secondary lobe structure of the power--pattern, and the sky at that instant. \label{fig:fhd_data}}
\end{figure}

@@

A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}. Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid. Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky, where the MWA tile power gain is significant ($\gtrsim 12$\%). In the {\it zenith} pointing, the Galactic plane has set and the contour level of the power--pattern in that direction is at least 16 times smaller.

@@

The model described in the preceding section is primarily a model of the diffuse foreground sky. While it contains faint compact emission blended in with the diffuse emission, it is known to be devoid of bright compact sources of emission \citep{deo08}. In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and use compact sources above this threshold. It may be noted that slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}. For an angular resolution of 0\fdg 85, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy. Other liberal criteria that yield a lower threshold carry a greater risk of double--counting compact sources which might be already blended in the diffuse sky model.

@@

It is worth emphasizing that the ideal foreground model is one with an angular resolution finer than that of the instrument that covers the whole sky at the observing frequency, a point we will revisit later for another reason. For this study, we note that our conclusions are not significantly affected despite minor differences in the thresholds from different radio source confusion estimates.

@@

Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of the compact foreground sky model (see Figure~\ref{fig:CSM}) for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels, respectively.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise. White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and logarithmic color scale are identical to those in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}. Emission from compact objects gives rise to centrally concentrated features and stays flat across antenna spacings in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
\end{figure}

In contrast to the delay spectrum of diffuse emission, compact emission manifests as a center--heavy structure in the delay spectrum with either pointing. This contributes to the {\it pitchfork} feature which will be discussed in detail in \S\ref{sec:composite}. Below we summarize some specific notable features obtained with our compact objects foreground model.

%\subsubsection{Compact Emission in Power--Pattern Main Lobe}\label{sec:csm-main-lobe}

@@

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as dotted, dashed, and solid lines, respectively. The horizon delay limits are shown as a pair of vertical dot--dashed lines. The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum. Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground objects, while thermal noise fluctuations dominate outside.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}). The antenna spacing is specified in each panel along East, North and Upward directions. The dotted, dashed and solid lines denote contributions from diffuse emission, compact objects and composite foreground model respectively. Vertical dot--dashed lines mark the horizon delay limits. The shaded region denotes the envelope of uncertainty around the composite model delay spectrum. This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits. Outside the horizon delay limits, it is predominantly due to thermal noise fluctuations. Compact objects dominate the central regions of delay spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure. \label{fig:pitchfork-baselines}}
\end{figure}

@@

The Galactic center in the {\it off--zenith} pointing is one such example already available in our study. Figure~\ref{fig:sky-model} shows the sky model ({\it top:} compact objects, {\it bottom:} diffuse sky model) in this pointing. The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co--located with a bright secondary lobe of the power pattern near the western horizon. Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented northward (top panel) and eastward (bottom panel). Since the Galactic center is located in the western sky, this figure demonstrates that it is observed at negative delay on a baseline vector oriented eastward and at zero delay on a baseline vector oriented northward. Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward (67\fdg 5~$\le \theta_\textrm{b} <$~112\fdg 5) at the top and eastward (-22\fdg 5~$\le \theta_\textrm{b} <$~22\fdg 5) at the bottom. The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}). Consequently, the spillover caused by the instrument's transfer function from the {\it foreground wedge} into the {\it EoR window} is most severe in eastward baselines (particularly the short ones) relative to northward baselines evident by the bright vertical stripes of foreground contamination.

\begin{figure*}[htb]
\centering
\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figures/v1_0/sky_model_snapshot_0.eps}}
\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figures/v1_0/directional_delay_map_snapshot_0.eps}}
\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
\caption{(a): Sky model showing compact objects ({\it top}) and diffuse emission ({\it bottom}). Contours of power pattern are overlaid whose levels are identical to those in Figures~\ref{fig:DSM} and \ref{fig:CSM}. The Galactic center is very prominent in diffuse emission in the western sky and is co--located with a bright secondary lobe of the power pattern. (b): Sky hemisphere mapped to delays observed on antenna spacings with northward ({\it top}) and eastward ({\it bottom}) orientations. Color scale shown is for a 100~m antenna spacing. Delays vary linearly with antenna spacing length. The bright Galactic center in the western sky is recorded at zero delays in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings. (c): Simulated delay spectra $|V_{b\tau}(\vec{b},\tau)|$ on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}). The bright Galactic center is prominently visible close to negative horizon delay limit, especially on short eastward antenna spacings. These are also the most severely contaminated by foreground spillover. The northward antenna spacings, on the other hand, are the least contaminated.}
\label{fig:breakup}
\end{figure*}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
\caption{Simulated delay spectra amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}). Discarded antenna spacings (black vertical stripes) have lengths $|\vec{b}| < 30$~m and orientations -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5. The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by an order of magnitude when short eastward antenna spacings are discarded. Color scale has been adjusted to emphasize mitigation achieved in foreground contamination. \label{fig:before-after}}
\end{figure}

This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data. Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{We use standard deviation of noiseless $V_{b\tau}(\vec{b},\tau)$ from foregrounds in the {\it EoR window} as a measure of foreground contamination.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings. We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars). Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\vec{b}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings. In other words, these plots demonstrate the progress in mitigation as orientation and maximum length of discarded antenna spacings are varied. Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits, respectively, and are shown for reference. The fraction of antenna spacings discarded relative to the total number of antenna spacings is shown in dotted lines for different ranges of $\theta_\textrm{b}$. It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5). It is notable that short antenna spacings play maximum role in determining foreground contamination. For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%. On the other hand, a bright compact object at the same location will give rise to foreground contamination even on longer antenna spacings. Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings accounting for $\sim 25$\% loss of data.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
\caption{Foreground contamination, estimated using standard deviation in the {\it EoR window}, for the {\it off--zenith} pointing as a function of orientation and length of discarded antenna spacings. Antenna spacings oriented eastward with varying degrees of directedness are considered: -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars). Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded (dotted lines). Lower (dashed line) and upper (dot--dashed) limits on foreground contamination are derived from northward (67\fdg 5~$\le\theta_\textrm{b}<$~112\fdg 5) and eastward ($|\theta_\textrm{b}|\le$~22\fdg 5) antenna spacings respectively. Foreground contamination in the {\it EoR window} (solid lines) reduces by factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5). Short baselines have maximum effect on foreground contamination. For instance, discarding antenna spacings oriented $|\theta_\textrm{b}|\le$~22\fdg 5 with lengths $|\vec{b}|\gtrsim 30$~m has no effect in further reducing foreground contamination. \label{fig:screening}}
\end{figure}

>>>>

Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}. Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity. Foregrounds can be described in both the delay spectrum and Fourier frameworks. For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite their subtle differences.

@@

Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper. $|\vec{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis. $\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by antenna spacing and the light travel time across it. These boundaries are called horizon delay limits and are shown by solid lines. The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis. This effect has been quantified in \citet{thy13}. The width of this narrow strip is inversely proportional to the operating bandwidth. The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades. However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.

@@

We show results of delay spectrum from MWA data after deconvolution along the delay axis in Figure~\ref{fig:fhd_data}. The principle feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time corresponding to baseline length. Note that the wedge is not filled out uniformly. Each baseline has a different orientation and resulting response to different objects in the sky. It is this variation that we will explore below.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{Delay spectra of two observations of the EoR field (RA = 0h, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set. For reference, the $k$--space equivalent of $\tau$ and $|\vec{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline. Smooth spectrum objects predominantly reside in this {\it foreground wedge}. The logarithmic color scale (shown at the right) is common to both panels. \label{fig:fhd_data}}
\end{figure}

@@

A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}. Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid. Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky, where the MWA tile power gain is significant ($\gtrsim 12$\%). In the {\it zenith} pointing, the Galactic plane has set and the power--pattern in that direction is at least 16 times smaller.

@@

The model described in the preceding section is primarily a model of the diffuse foreground sky. While it contains faint compact emission blended in with the diffuse emission, it is known to be devoid of bright compact sources \citep{deo08}. In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and use compact sources above this threshold. It may be noted that slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}. For an angular resolution of 0\fdg 85, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy. Other liberal criteria that yield a lower threshold carry a greater risk of double--counting compact sources which might be already blended in with the diffuse sky model.

@@

It is worth emphasizing that the ideal foreground model is one with an angular resolution finer than that of the instrument that covers the whole sky at the observing frequency, a point we will revisit later. For this study, we note that our conclusions are not significantly affected despite minor differences in the thresholds from different radio source confusion estimates.

@@

Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of the compact foreground sky model (see Figure~\ref{fig:CSM}) for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels, respectively. In contrast to the delay spectrum of diffuse emission, compact emission manifests as a center--heavy structure in the delay spectrum with either pointing.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Same as Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a foreground model consisting of compact emission. Compact emission gives rise to centrally concentrated features and stays flat across antenna spacings in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
\end{figure}

@@

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum}, and \ref{fig:noiseless-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively. The horizon delay limits are shown as a pair of vertical dotted lines. The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum. Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground objects, while thermal noise fluctuations dominate outside.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}). The antenna spacing is specified in each panel along East, North and Upward directions. The solid red, cyan, and black lines denote contributions from diffuse emission, compact objects and composite foreground model respectively. Vertical dotted lines mark the horizon delay limits. The shaded region denotes the envelope of uncertainty around the composite model delay spectrum. This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits. Outside the horizon delay limits, it is predominantly due to thermal noise fluctuations. Compact objects dominate the central regions of delay spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure. \label{fig:pitchfork-baselines}}
\end{figure}

@@

The Galactic center in the {\it off--zenith} pointing is one such example already available in our study. Figure~\ref{fig:sky-model} shows the sky model ({\it top:} compact objects, {\it bottom:} diffuse sky model) in this pointing. The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co--located with a bright secondary lobe of the power pattern near the western horizon. Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented northward (top panel) and eastward (bottom panel). Since the Galactic center is located in the western sky, this figure demonstrates that it is observed at negative delay on a baseline vector oriented eastward and at zero delay on a baseline vector oriented northward. Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward (67\fdg 5~$\le \theta_\textrm{b} <$~112\fdg 5) at the top and eastward (-22\fdg 5~$\le \theta_\textrm{b} <$~22\fdg 5) at the bottom. The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}). Consequently, the spillover caused by the instrument's transfer function from the {\it foreground wedge} into the {\it EoR window} affects the northward baselines the least and is most severe in eastward baselines (particularly the short ones) evident by the bright vertical stripes of foreground contamination.

\begin{figure*}[htb]
\centering
\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figures/v1_0/sky_model_snapshot_0.eps}}
\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figures/v1_0/directional_delay_map_snapshot_0.eps}}
\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
\caption{(a): Sky model showing compact objects ({\it top}) and diffuse emission ({\it bottom}). Contours of power pattern are overlaid whose levels are identical to those in Figures~\ref{fig:DSM} and \ref{fig:CSM}. The Galactic center is very prominent in diffuse emission in the western sky and is co--located with a bright secondary lobe of the power pattern. (b): Sky hemisphere mapped to delays observed on antenna spacings with northward ({\it top}) and eastward ({\it bottom}) orientations. Color scale shown is for a 100~m antenna spacing. Delays vary linearly with antenna spacing length. The bright Galactic center in the western sky is recorded at zero delays in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings. (c): Simulated delay spectra $|V_{b\tau}(\vec{b},\tau)|$ on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}). White lines denote horizon delay limits. The bright Galactic center is prominently visible close to negative horizon delay limit, especially on short eastward antenna spacings. These are also the most severely contaminated by foreground spillover. The northward antenna spacings, on the other hand, are the least contaminated.}
\label{fig:breakup}
\end{figure*}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
\caption{Simulated delay spectra amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}). Discarded antenna spacings (black vertical stripes) have lengths $|\vec{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22\fdg 5. The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by an order of magnitude when short eastward antenna spacings are discarded. Color scale has been adjusted to emphasize mitigation achieved in foreground contamination. \label{fig:before-after}}
\end{figure}

This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data. Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{We use standard deviation of noiseless $V_{b\tau}(\vec{b},\tau)$ from foregrounds in the {\it EoR window} as a measure of foreground contamination.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings. We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars). Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\vec{b}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings. In other words, these plots demonstrate the progress in mitigation as orientation and maximum length of discarded antenna spacings are varied. % Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits, respectively, and are shown for reference. 
The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$. It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5). It is notable that short antenna spacings play maximum role in determining foreground contamination. For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%. On the other hand, a bright compact object at the same location will give rise to foreground contamination even on longer antenna spacings. Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings accounting for $\sim 25$\% loss of data.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
\caption{Foreground contamination, estimated using standard deviation in the {\it EoR window}, for the {\it off--zenith} pointing as a function of orientation and length of discarded antenna spacings. Antenna spacings oriented eastward with varying degrees of directedness are considered: -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars). Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded. The fraction of discarded antenna spacings relative to the total number is denoted by dotted lines for the corresponding cases. % Lower (dashed line) and upper (dot--dashed) limits on foreground contamination are derived from northward (67\fdg 5~$\le\theta_\textrm{b}<$~112\fdg 5) and eastward ($|\theta_\textrm{b}|\le$~22\fdg 5) antenna spacings respectively. 
  Foreground contamination in the {\it EoR window} (solid lines) reduces by factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5). Short baselines have maximum effect on foreground contamination. For instance, discarding antenna spacings oriented $|\theta_\textrm{b}|\le$~22\fdg 5 with lengths $|\vec{b}|\gtrsim 30$~m has no effect in further reducing foreground contamination. \label{fig:screening}}
\end{figure}
